|

### 2021-12-06 06-Chp-IPD-Analysis.Rmd
End of day:

Individual participant data (IPD) meta-analyses are a special type of meta-analysis that contact the authors of relevant studies to request access to the raw data on each participant. This is different to the meta-analyses presented in the previous chapter, which used published summary results.

@@

To examine the relationship between lipid levels and dementia risk, I applied for access to `r n_applied$total` unique data sources containing information on these variables. However, only a small proportion of these data sources (`r fmt_applied(n_accessed$total)`) provided me with access.

@@

The resulting analysis of these data sources did not suggest a relationship between any blood lipid and any dementia outcome. The sole exception was an increased risk of vascular dementia in those with higher triglyceride levels. In addition, participants age and sex did not appear to influence the relationship of lipid fractions and dementia outcomes.

@@

The reasons for the low response rate to request for data are explored. Finally, I suggest future studies to investigate how data access rates could be improved.

@@

Individual participant data (IPD) meta-analyses are considered to be the gold standard form of evidence synthesis, allowing for the application of a common selection model and analytical approach across all identified cohorts.[@riley2010] They are particularly useful for investigating the impact of participant-level characteristics, something that is not possible with aggregate data unless the results are stratified by the characteristic of interest.[@riley2010;@thompson2005] Knowledge of which groups a treatment will benefit most (or harm least) is a core aim of the move towards personalised medicine.[@riley2020;@hingorani2013]
IPD analysis also offer a mechanism by which previously unanalysed datasets can be incorporated into an analysis, thus expanding the evidence base for a particular research question.

@@

Eligible data sources for this analysis were prospective cohort studies. Data sources which were cross-sectional, either by design or due to the available data (e.g. a study conducted across multiple waves, but only data from a single wave could be accessed) were excluded. Similarly, due to time restrictions within the scope of my thesis project, studies making use of population-level electronic health records, which often require an extensive project proposals in order to gain access to the data, were ineligible due to the time and cost involved in applying. No restrictions were put on the number of participants or the length of follow-up, though participants must be free of dementia (or assumed to be free, based on age at entry)  at baseline.

@@

Potentially eligible data sources as defined above were identified via two mechanisms, each with a distinct focus. For both approaches, the number of cohorts responding to the request for data access, and where applicable, the reasons given for a lack of data sharing were recorded.

@@

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapter \@ref(sys-rev-heading)  The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the text and documentation sent to potentially collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who has moved to a different institution.

@@

Where data on one of these exposures (TC, LDL-c, HDL-c, TG) was missing, it was inferred from the other three fractions (where available) using the Friedwald formula (Equation \@ref(eq:total-cholesterol-formula)). Lipids levels were retained as continuous variables rather than dichotomising into a binary hypercholesterolemia exposure, given the additional power this adds to the meta-analysis.[@riley2020;@ensor2018] Where lipid measurements were reported in _mg/dL_, these were converted to _mmol/L_.

@@

Across all cohorts, data cleaning was performed in a similar manner, standardising to commonly named variables so that a single model could be applied using functional programming.[@wickham2016func] The advantage of this approach is that it reduces the likelihood of errors in model mis-specification if needing to change variables names from cohort to cohort. Following data cleaning, summary statistics for each data source were calculated and compared with publicly available statistics to ensure no errors were introduced in the data cleaning process, in line with best practice.[@levis2021]

@@

In order to investigate the interaction of participant-level characteristics (age and sex) with lipid levels, interaction terms for lipid-covariate terms were extracted and synthesised using a fixed effects meta-analysis.[@fisher2017] To avoid ecological bias, where the between-study association does not reflect the within-study associations, cohorts where there was no within-study variation in the covariate of interest were excluded from the interaction analysis for that covariate.[@burke2017] As an example, it is impossible to estimate the impact of sex on the lipid~dementia relationship in a study which contains female participants. All interaction analysis were performed using the common-set-adjusted model (Model 1).

@@

Given minimal effect of further adjustment for education and BMI on the effect estimates, the results from the interaction analysis presented here were obtained using the common-set-adjusted model. Further, while the maximally-adjusted could have been employed in the analysis of the effect of sex, as the CaPS cohort contains only men and had to be excluded from this analysis, employing the same underlying model helped to ensure comparability between the age and sex interaction estimates.

@@

Similar to the analysis of CPRD data presented in the previous chapter, this discussion will not provide a detailed comparison of the results of this analysis with other published literature, except to compare between this analysis and previously published results using the same data source. For a detailed comparison of the findings presented above with the existing evidence base identified by the systematic review (Chapter \@ref(sys-rev-heading)), see Chapter \@ref(tri-heading).

@@

@@

There are likely several likely reasons for this low response rate to requests for data access. In general terms, there are a several well-documented barriers that prevent data from being made readily available, including concerns regarding participant privacy, fear of "scooping" or "parasitic" behaviour, and a lack of trust between primary and secondary researchers.[@vanpanhuis2014] More specific to this analysis, IPD meta-analysis including studies other than randomised controlled trials have less success in obtaining IPD from studies.[@nevitt2017] Additionally, while no evidence is available on whether the characteristics of the researcher requesting data access influences the response rate and eventual decision, there is the possibility that my position as a PhD student meant I was less likely to elicit the response rate than a well-known senior academic might. The timing of the requests for data access, coinciding with a global pandemic, may also have affected the response rate as researchers prioritised COVID-related work.

@@

While this analysis did not manage to systematically obtain and analyse a large proportion of identified data, a central strength of this analysis is use of a systematic approach to identify and attempt to contact relevant cohorts. Further, it also enabled incorporation of two previously unanalysed datasets - the CaPS and EPIC Norfolk cohorts - thus providing additional evidence that is used in the triangulation exercise reported in Chapter \@ref(tri-heading).

A further strength is the ability of this analysis was the ability to investigate the effect of participant characteristics (namely age and sex) on the association of blood lipids and dementia outcomes. To the best of my knowledge, no previous review has examined the interaction of these factors with the observed associations.

@@

@@

- In this chapter, I performed an IPD meta-analysis to investigate the effect of blood lipid levels on risk of incident dementia. There was a very low response rate to requests for data access, resulting in the inclusion of only three cohorts.

@@

- I discussed potential reasons for the low response rate, and explored other limitations of this analysis. I highlighted the contribution of this work to the wider topic via the analysis of previously unexplored cohorts (CaPS & EPIC) and outcomes (vascular dementia). Finally, I recommended that future research formally investigate the impact of request characteristics on requests for data.

>>>>

Individual participant data (IPD) meta-analyses are a special type of meta-analysis where those undertaking the study contact the authors of relevant studies to request access to the raw data on each participant. This is different to the meta-analyses presented in the previous chapter, which used published summary results.

@@

To examine the relationship between lipid levels and dementia risk, I applied for access to `r n_applied$total` unique data sources containing information on these variables. However, only a small proportion of these data sources (`r fmt_applied(n_accessed$total)`) provided the requested access.

@@

The resulting analysis of these data sources did not suggest a relationship between any blood lipid and any dementia outcome. The sole exception was an increased risk of vascular dementia in those with higher triglyceride levels. In addition, the participants age and sex did not appear to influence the relationship of lipid fractions and dementia outcomes.

@@

The reasons for the low response rate to requests for data are explored. Finally, I suggest future studies to investigate how data access rates could be improved.

@@

Individual participant data (IPD) meta-analyses are considered to be the gold standard form of evidence synthesis, allowing for the application of a common selection model and analytical approach across all identified cohorts.[@riley2010] They are particularly useful when investigating the impact of participant-level characteristics, something that is not possible with aggregate data unless the results are stratified by the characteristic of interest.[@riley2010;@thompson2005] Knowledge of which groups a treatment will benefit most (or harm least) is a core aim of the move towards personalised medicine.[@riley2020;@hingorani2013]
IPD analysis also offer a mechanism by which previously unanalysed datasets can be incorporated into an analysis, thus expanding the evidence base for a particular research question.

@@

Eligible data sources for this analysis were prospective cohort studies. Data sources which were cross-sectional, either by design or due to the available data (e.g. a study conducted across multiple waves, but only data from a single wave could be accessed) were excluded. Similarly, studies making use of population-level electronic health records, which often require an extensive project proposals in order to gain access to the data, were ineligible due to the time and cost involved in applying. No restrictions were put on the number of participants or the length of follow-up, though participants must be free of dementia (or assumed to be free, based on age at entry)  at baseline.

@@

Potentially eligible data sources as defined above were identified via two mechanisms, each with a distinct focus. For both approaches, the number of cohorts responding to the request for data access, and where applicable, the reasons given for a refusal were recorded.

@@

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapter \@ref(sys-rev-heading)  The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the text and documentation sent to potential collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

Where data on one of these exposures (TC, LDL-c, HDL-c, TG) was missing, it was inferred from the other three fractions (where available) using the Friedwald formula (Equation \@ref(eq:total-cholesterol-formula)). Lipids levels were retained as continuous variables rather than dichotomising into a binary hypercholesterolemia exposure, given the additional statistical power this adds to the meta-analysis.[@riley2020;@ensor2018] Where lipid measurements were reported in _mg/dL_, these were converted to _mmol/L_.

@@

Across all cohorts, data cleaning was performed in a similar manner, standardizing to commonly named variables so that a single model could be applied using functional programming.[@wickham2016func] The advantage of this approach is that it reduces the likelihood of errors in model mis-specification if changes are required in variables names from cohort to cohort. Following data cleaning, summary statistics for each data source were calculated and compared with publicly available statistics to ensure no errors were introduced in the data cleaning process, in line with best practice.[@levis2021]

@@

In order to investigate the interaction of participant-level characteristics (age and sex) with lipid levels, interaction terms for lipid-covariate terms were extracted and synthesised using a fixed effects meta-analysis.[@fisher2017] To avoid ecological bias, where the between-study association does not reflect the within-study associations, cohorts where there was no within-study variation in the covariate of interest were excluded from the interaction analysis for that covariate.[@burke2017] As an example, it is impossible to estimate the impact of sex on the lipid~dementia relationship in a study which contains only female participants. All interaction analysis were performed using the common-set-adjusted model (Model 1).

@@

Given minimal effect of further adjustment for education and BMI on the effect estimates, the results from the interaction analysis presented here were obtained using the common-set-adjusted model. Further, while the maximally-adjusted model (Model 2) could have been employed in the analysis of the effect of sex, as the CaPS cohort contains only men and had to be excluded from this analysis, employing the same underlying model helped to ensure comparability between the age and sex interaction estimates.

@@

A detailed comparison of the findings presented above with the existing evidence base identified by the systematic review (Chapter \@ref(sys-rev-heading)), is presented as part of the triangulation exercise in Chapter \@ref(tri-heading). As such, this discussion will not provide a detailed comparison of the results of this analysis with other published literature, except to compare between this analysis and previously published results using the same data source.

@@

<!-- REVIEW also identified that any future studies would require a longer and/or more directive approach to the provision of access - DPUK is a good example of how this could be done! -->

@@

There are likely several likely reasons for this low response rate to requests for data access. In general terms, there are a several well-documented barriers that prevent data from being made readily available, including concerns regarding participant privacy, fear of "scooping" or "parasitic" behaviour, and a lack of trust between primary and secondary researchers.[@vanpanhuis2014] More specific to this analysis, IPD meta-analysis including studies other than randomised controlled trials have less success in obtaining IPD from studies.[@nevitt2017] Additionally, while no evidence is available on whether the characteristics of the researcher requesting data access influences the response rate and eventual decision, there is the possibility that my position as a PhD student meant I was less likely to elicit a positive response than a well-known senior academic might. The timing of the requests for data access, coinciding with a global pandemic, may also have affected the response rate as researchers prioritised COVID-related work.

@@

While this analysis did not manage to systematically obtain and analyse a large proportion of identified data, a central strength of this analysis is the use of a systematic approach to identify and attempt to contact relevant cohorts. Further, it also enabled incorporation of two previously unanalysed datasets - the CaPS and EPIC Norfolk cohorts - thus providing additional evidence that is used in the triangulation exercise reported in Chapter \@ref(tri-heading).

A further strength of this analysis was the ability to investigate the effect of participant characteristics (namely age and sex) on the association of blood lipids and dementia outcomes. To the best of my knowledge, no previous review has examined the interaction of these factors with the observed associations.

@@

<!-- REVIEW Or a requirement of peer review and publication be that the data be recorded centrally and made available through a coordinated application process) -->

@@

- In this chapter, I performed an IPD meta-analysis to investigate the effect of blood lipid levels on the risk of incident dementia. There was a very low response rate to requests for data access, resulting in the analysis of three relevant cohorts.

@@

- I discussed potential reasons for the low response rate, and explored other limitations of this analysis. I highlighted the contribution of this work to the wider topic via the analysis of previously unexplored cohorts (CaPS & EPIC) and outcomes (vascular dementia). Finally, I recommended that future research formally investigate the impact of requestor characteristics on requests for data.

### 2021-12-06 07-Chp-Triangulation.Rmd
End of day:

@@

#### Defining reasonable prior distributions for bias and indirectness

>>>>

<!-- TODO Sensitivity analysis using only heterogeneity - no change to point estimate - places less importance on the accurate determination of the direction of bias -->

@@

#### Defining reasonable prior distributions for bias and indirectness {#tri-prior-def}
